The {TestIdeals} package for \emph{Macaulay2}

### 2018-10-04 Fsing/TestIdealsPaper/TestIdeals.tex
Small changes to paper

\section{Ideals compatible with given $p^{-e}$-linear map}\label{Section: compatible ideals}

Throughout this section let $R$ denote a polynomial $\mathbb{K}[x_1, \dots, x_n]$. In this section we address the following question:
given a $R$-linear map $\phi: R^{1/p^e} \rightarrow R$, what are all ideals $I\subseteq R$ such that $\phi(I^{1/p^e})=I$?
We call these ideals $\phi$-compatible.

Recall that  $\Hom_R( R^{1/p^e}, R)$  is a principal $R^{1/p^e}$-module generated
the \emph{trace map} $T\in \Hom_R(R^{1/p^e}, R)$, constructed as follows (cf. \cite[Lemma 1.6]{FedderFPureRat} and \cite[Example 1.3.1]{BrionKumarFrobeniusSplitting}).
Recall that, if $B$ is a $\mathbb{K}$-basis for $\mathbb{K}^{1/p^e}$ (with $1\in B$),
$R^{1/p^e}$ is a free $R$-module with basis
$$\left\{ b x_1^{\alpha_1/p^e} \dots x_n^{\alpha_n/p^e} \,|\, 0\leq \alpha_1, \dots, \alpha_n < p^e \right\} ;$$
the trace map $T$ is the projection onto the free summand
$R x_1^{(p^e-1)/p^e} \cdots x_n^{(p^e-1)/p^e}\cong R$.

@@

The method \emph{compatibleIdeals} produces the finite set of $\phi$-compatible prime ideals in the second statement of Theorem \ref{Theorem: finitely many compatible primes}.

@@

The defining condition $u I = I^{[p]}$ for $\phi$-compatible ideals allows us to
think of these in a dual form: write $\mathfrak{m}=(x_1, \dots, x_n)R$,
$E=E_{R_\mathfrak{m}}(R_{\mathfrak{m}}/\mathfrak{m})=H^n_{\mathfrak{m}} (R)$, and let
$\Theta: E \rightarrow E$ be the $p^e$-linear map\footnote{I.e., $\Theta$ is additivive and $\Theta (r a)= r^{p^e} \Theta (a)$ for all $a\in E$ and $r\in R$.}
induced from the Frobenius map on $R$.
If $\psi=u \Theta$, then $\psi \Ann_E I \subseteq \Ann_E I$ if and only if $u I = I^{[p]}$!
Thus finding all $R$-submodules of $E$ compatible with $\psi=u \Theta$ also amounts to finding all
$\phi=T \circ u^{1/p^e}$-compatible ideals.

\begin{example}

We return to Example \ref{Example: parameter test ideal}.
In (cf.~\cite[\S 9]{KatzmanParameterTestIdealOfCMRings}) it is shown that
there is a surjection $\Ann_E I \rightarrow H^2_{\mathfrak{m}} (R/I)$
which is compatible with with the induced $p^1$-linear map on $H^2_{\mathfrak{m}} (R/I)$
and the $p^1$-linear map $u \Theta$ on $\Ann_E I$ where $u$ is computed as follows.

@@

In \cite{KatzmanParameterTestIdealOfCMRings} it is shown that as that $R/I$ is Cohen-Macaulay $u$
can be taken as the generator of the cyclic module $(I^{[p]}:I) \cap (\Omega^{[p]}:\Omega)/I^{[p]}$
where $\Omega \subseteq R$ is an ideal whose image in $I$ is a canonical module of $R/I$.

@@

Now we can compute all annihilators of $R$-submodules of $E$ stable under the $p^1$-linear map $uT$

@@

and all annihilators of $R$-submodules of $H^2_{\mathfrak{m}} (R/I)$ stable under the $p^1$-linear map induced from the Frobenius map on $R/I$.

@@

In \cite{KatzmanZhangAlgorithm} the algorithms behind the method in Section \ref{Section: compatible ideals} were extended
to compute prime annihilators of submodules of Artinian modules compatible with a given $p^{e}$-linear map.
This would require, among other things, a faster implementation of our method for finding Frobenius roots of submodules of free modules.

>>>>

\section{Ideals compatible with a given $p^{-e}$-linear map}\label{Section: compatible ideals}

Throughout this section, let $R$ denote a polynomial $\mathbb{K}[x_1, \dots, x_n]$.
In this section we address the following question:
given a $R$-linear map $\phi: R^{1/p^e} \rightarrow R$, what are the ideals $I\subseteq R$ such that $\phi(I^{1/p^e})\subseteq I$?
We call these ideals $\phi$-compatible.

Recall that  $\Hom_R( R^{1/p^e}, R)$  is a principal $R^{1/p^e}$-module generated
the \emph{trace map} $T\in \Hom_R(R^{1/p^e}, R)$, constructed as follows (cf. \cite[Lemma 1.6]{FedderFPureRat} and \cite[Example 1.3.1]{BrionKumarFrobeniusSplitting}).
If $B$ is a $\mathbb{K}$-basis for $\mathbb{K}^{1/p^e}$, with $1\in B$, then 
$R^{1/p^e}$ is a free $R$-module with basis
$$\left\{ b\,  x_1^{\alpha_1/p^e} \cdots x_n^{\alpha_n/p^e} \,|\, 0\leq \alpha_1, \dots, \alpha_n < p^e, b\in B \right\} ;$$
the trace map $T$ is the projection onto the free summand
$R x_1^{(p^e-1)/p^e} \cdots x_n^{(p^e-1)/p^e}\cong R$.
%
\pedro{This paragraph is a bit redundant; we could just refer to Section~\ref{ss: math behind}}
%

@@

The method \texttt{compatibleIdeals} produces the finite set of $\phi$-compatible prime ideals in the second claim of Theorem \ref{Theorem: finitely many compatible primes}.

@@

The defining condition $u I \subseteq I^{[p^e]}$ for $\phi$-compatible ideals allows us to
think of these in a dual form: write $\mathfrak{m}=(x_1, \dots, x_n)$,
$E=E_{R_\mathfrak{m}}(R_{\mathfrak{m}}/\mathfrak{m})=H^n_{\mathfrak{m}} (R)$, and let
$\Theta: E \rightarrow E$ be the $p^e$-linear map\footnote{That is, $\Theta$ is additivive and $\Theta (r a)= r^{p^e} \Theta (a)$ for all $a\in E$ and $r\in R$.}
induced from the Frobenius map on $R$.
If $\psi=u \Theta$, then $\psi \Ann_E I \subseteq \Ann_E I$ if and only if $u I \subseteq  I^{[p^e]}$.
Thus finding all $R$-submodules of $E$ compatible with $\psi=u \Theta$ also amounts to finding all
$\phi$-compatible ideals, where $\phi=T \circ u^{1/p^e}$.

\begin{example}
We return to Example \ref{Example: parameter test ideal}.
In (cf.~\cite[\S 9]{KatzmanParameterTestIdealOfCMRings}) it is shown that
there is a surjection $\Ann_E I \rightarrow H^2_{\mathfrak{m}} (R/I)$
which is compatible  with the induced $p^1$-linear map on $H^2_{\mathfrak{m}} (R/I)$
and the $p^1$-linear map $u \Theta$ on $\Ann_E I$, where $u$ is computed as follows.

@@

In \cite{KatzmanParameterTestIdealOfCMRings} it is shown that as $R/I$ is Cohen-Macaulay, $u$
can be taken as the generator of the cyclic module $(I^{[p]}:I) \cap (\Omega^{[p]}:\Omega)/I^{[p]}$
where $\Omega \subseteq R$ is an ideal whose image in $R/I$ is a canonical module of that ring.

@@

Now we can compute all annihilators of $R$-submodules of $E$ stable under the $p^1$-linear map $uT$.

@@

We can also compute  all annihilators of $R$-submodules of $H^2_{\mathfrak{m}} (R/I)$ stable under the $p^1$-linear map induced from the Frobenius map on $R/I$.

@@

In \cite{KatzmanZhangAlgorithm}, the algorithms behind the methods in Section \ref{Section: compatible ideals} were extended
to compute prime annihilators of submodules of Artinian modules compatible with a given $p^{e}$-linear map.
This would require, among other things, a faster implementation of our method for finding Frobenius roots of submodules of free modules.
